---
title: "HW3 WH Salaries"
---

In this assignment, we will be working with a dataset containing information about the salaries of White House staffers as of July 2026.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2026.csv. You will create a dataframe called `wh_salaries` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
wh_salaries <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2026.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, what stands out to you about this data? Are there any salary values that will cause problems for analysis?

#The chart shows the name and status of each employee, as well as their job title, salary, and basis of pay. From an analysis standpoint, several employees have a salary of $0 listed. I'm not sure whether this is intentional or no figures were to be found, bute either way it might present problems.

### Task 4: Calculate the Mean and Median of the Salaries

Write code to calculate the mean and median of the salaries in the dataset using the `summarize()` function. You don't need to make a new dataframe, just display the results.

```{r}
wh_salaries |> 
  summarize(
    mean_salary = mean(SALARY),
    median_salary =median(SALARY)
  )
```

### Task 5: Add in the minimum and maximum salaries

Take the code you just wrote and add in the minimum and maximum salaries (the functions for that are `min()` and `max()`). Your results should show four columns: mean_salary, median_salary, min_salary, max_salary.

```{r}
wh_salaries |> 
  summarize(
    mean_salary = mean(SALARY),
    median_salary =median(SALARY),
    min_salary =min(SALARY),
    max_salary =max(SALARY)
  )
```

### Task 6: Describe the Results

Write a couple of sentences describing the results of your analysis. What do the mean and median salaries tell you about the data? What do the minimum and maximum salaries tell you? Are there any outliers in the data?

#The mean salary is around $112,702.70 and the median is $109,243. The minimus is $0 and the maximus is $197,000. I think the data definitely has outliers, as seen by the values of 0 entered in the chart. These numbers are definitely skewing the data, and I would be curious to see the average salary without them.

### Task 7: Story Idea

Based on the data, what is your best story idea? What is the next step you would take to develop that story idea?

#I think my best story idea woulde center the missing salaries. Why are they missing? What do those people do? I would try contacting them and people who work with them in order to see if I could find the values or see why not.
